Constructive Preference Elicitation by Setwise Max-margin Learning

### 2016-01-28 paper/ijcai16.tex
paper: 9points for captions

\documentclass{article}
\usepackage{ijcai16}
\usepackage{times}
%\usepackage{latexsym}
%\usepackage{graphicx}
\usepackage{graphbox}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{placeins}

@@

Consistently with previous work~\cite{guo2010real,viappiani2010optimal}, 
we model users with additive utility functions \cite{keeney1976}; 
the user's preferences are represented by a  weight
vector $\vw\in\bbR^m$ and
%We assume linear utility functions \cite{keeney1976}; 
the utility of a configuration $\vx$ is given by
$\langle \vw, \vx \rangle = \sum_{z=1}^m w_z x_z$. 
In the remainder of the paper we
require all weights to be {\em non-negative} and {\em bounded}: the per-attribute
weights $w_z$ must lie in a (constant but otherwise arbitrary) interval
$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. 
Both requirements are quite natural\footnote{Utility values are defined on an
interval scale, therefore the assumption of non negativity is not restrictive as it is always possible to scale the values appropriately (see for instance \cite{Torra2007} and \cite{keeney1976}).}, and enable the translation of our core optimization problem into a
mixed-integer linear problem (as done in Section~\ref{sec:formulation}).

@@

We compare {\sc SetMargin} against three state-of-the-art Bayesian
approaches: i) the Bayesian approach from \cite{guo2010real},
selecting queries according to {\em restricted informed VOI} ({\sc
  riVOI}), a computationally efficient heuristic approximation of
value-of-information, and inference using TrueSkill$^{TM}$
\cite{HerbrichMG06} (an inference scheme based on expectation
propagation \cite{Minka01}); ii) the Bayesian framework of
\cite{viappiani2010optimal} using Monte Carlo methods \footnote{We use
  50000 particles to represent the probability distribution.}for
Bayesian inference and asking choice queries (asking the user to
select its preferred item among a set of diplayed ones); the next
query is computed by performing a greedy optimization of {\em Expected
  Utility of a Selection} (hereafter just called {\sc EUS}), a
submodular quantity that can be used as a proxy for value of
information (giving strong approximate guarantees); iii) {\em Query
  Iteration} (referred as {\sc QI} below), also from
\cite{viappiani2010optimal}, an even faster query selection method
based on sampling a set of utility vectors, selecting the items that
maximize utility for each of the vector and forming a first proposal
for a query set. Then, a new set is computed by considering the
optimal items in the associated beliefs, and the process is repeated
until convergence.

@@

\paragraph{Synthetic Dataset.} Following the experimental protocol
in~\cite{guo2010real} and \cite{viappiani2010optimal}, in the first
experiment we evaluate the behavior of the proposed method in an
artificial setting with increasingly complex problems. We developed
synthetic datasets with $r$ attributes, for increasing values of $r$.
Each attribute takes one of $r$ possible values, so that the one-hot
encoding of attributes results in $m=r^2$ features. In terms of space
of configurations, for $r=3$ the synthetic dataset corresponds to 
$\calX_\text{feasible} = [3] \times [3] \times [3]$, for $r=4$ to
$\calX_\text{feasible} = [4] \times [4] \times [4] \times [4]$, and so on. The
cardinality of $\calX_\text{feasible}$ is $r^r$, and grows (super)
exponentially with $r$.
% We developed
% seven synthetic datasets: each dataset involves $m=2,\ldots,7$
% attributes, where each attribute takes one of $m$ possible
% values. More explicitly, for $m=2$ the synthetic dataset is
% $\calX_\text{feasible} = [2] \times [2]$, for $m=3$ to
% $\calX_\text{feasible} = [3] \times [3] \times [3]$, and so on. The
% cardinality of $\calX_\text{feasible}$ is $m^m$, and grows (super)
% exponentially with $m$. 
For $r=3$, the dataset is comparable in size to the synthetic one used
in~\cite{guo2010real} and~\cite{viappiani2010optimal}.
%, which have
%three attributes with 2, 2 and 5 values respectively, for a total of
%20 feasible configurations.  
For larger $r$ the size of the space grows much larger than the one
typically used in the Bayesian preference elicitation literature, and
as such represents a good testbed for comparing the scalability of the
various methods. The feasible configuration space was encoded in {\sc
  SetMargin} through appropriate MILP constraints, while the other
methods require all datasets to be explicitly grounded.  Users were
simulated by drawing $20$ random utility vectors from each of four
different distributions. The first two mimic those used
in~\cite{guo2010real}\footnote{The normal distribution reported
  in~\cite{guo2010real} differs slightly from the one implemented in
  the code of their experimental setting.  In our experiments we use
  the latter.}: (1) a uniform distribution over $[1, 100]$ for each
individual weight, and (2) a normal distribution with mean $25$
and %covariance
standard deviation $\frac{25}{3}$ (each attribute is sample i.i.d).
We further produced two novel {\em sparse} versions of the uniform and
normal distributions setting to zero $80\%$ of the entries (sampled
uniformly at random). We set a maximum budget of 100 iterations for all
methods for simplicity.

@@

\paolo{any idea about possible future works?}
% FUTURE WORKS

>>>>

\documentclass{article}
\usepackage{ijcai16}
\usepackage{times}
%\usepackage{latexsym}
%\usepackage{graphicx}
\usepackage{graphbox}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{placeins}
\usepackage[font=small]{caption}

@@

Consistently with previous work~\cite{guo2010real,viappiani2010optimal}, 
we model users with additive utility functions \cite{keeney1976}; 
the user's preferences are represented by a  weight
vector $\vw\in\bbR^m$ and
%We assume linear utility functions \cite{keeney1976}; 
the utility of a configuration $\vx$ is given by
$\langle \vw, \vx \rangle = \sum_{z=1}^m w_z x_z$. 
In the remainder of the paper we
require all weights to be {\em non-negative} and {\em bounded}: the per-attribute
weights $w_z$ must lie in a (constant but otherwise arbitrary) interval
$[w^\bot_z, w^\top_z]$, with $w^\bot_z \ge 0$. 
Both requirements are quite natural\footnote{Utility values are defined on an
interval scale, thus it is always possible to scale the values appropriately (see for instance \cite{Torra2007} and \cite{keeney1976}).}, and enable the translation of our core optimization problem into a
mixed-integer linear problem (as done in Section~\ref{sec:formulation}).

@@

We compare {\sc SetMargin} against three state-of-the-art Bayesian
approaches: i) the Bayesian approach from \cite{guo2010real},
selecting queries according to {\em restricted informed VOI} ({\sc
  riVOI}), a computationally efficient heuristic approximation of
value-of-information, and inference using TrueSkill$^{TM}$
\cite{HerbrichMG06} (an inference scheme based on expectation
propagation \cite{Minka01}); ii) the Bayesian framework of
\cite{viappiani2010optimal} using Monte Carlo methods (with 50,000 particles) for
Bayesian inference and asking choice queries (asking the user to
select its preferred item among a set of diplayed ones); the next
query is computed by performing a greedy optimization of {\em Expected
  Utility of a Selection} (hereafter just called {\sc EUS}), a
submodular quantity that can be used as a proxy for value of
information (giving strong approximate guarantees); iii) {\em Query
  Iteration} (referred as {\sc QI} below), also from
\cite{viappiani2010optimal}, an even faster query selection method
based on sampling a set of utility vectors, selecting the items that
maximize utility for each of the vector and forming a first proposal
for a query set. Then, a new set is computed by considering the
optimal items in the associated beliefs, and the process is repeated
until convergence.

@@

\paragraph{Synthetic Dataset.} Following the experimental protocol
in~\cite{guo2010real} and \cite{viappiani2010optimal}, in the first
experiment we evaluate the behavior of the proposed method in an
artificial setting with increasingly complex problems. We developed
synthetic datasets with $r$ attributes, for increasing values of $r$.
Each attribute takes one of $r$ possible values, so that the one-hot
encoding of attributes results in $m=r^2$ features. In terms of space
of configurations, for $r=3$ the synthetic dataset corresponds to 
$\calX_\text{feasible} = [3] \times [3] \times [3]$, for $r=4$ to
$\calX_\text{feasible} = [4] \times [4] \times [4] \times [4]$, and so on. The
cardinality of $\calX_\text{feasible}$ is $r^r$, and grows (super)
exponentially with $r$.
% We developed
% seven synthetic datasets: each dataset involves $m=2,\ldots,7$
% attributes, where each attribute takes one of $m$ possible
% values. More explicitly, for $m=2$ the synthetic dataset is
% $\calX_\text{feasible} = [2] \times [2]$, for $m=3$ to
% $\calX_\text{feasible} = [3] \times [3] \times [3]$, and so on. The
% cardinality of $\calX_\text{feasible}$ is $m^m$, and grows (super)
% exponentially with $m$. 
For $r=3$, the dataset is comparable in size to the synthetic one used
in~\cite{guo2010real} and~\cite{viappiani2010optimal}.
%, which have
%three attributes with 2, 2 and 5 values respectively, for a total of
%20 feasible configurations.  
For larger $r$ the size of the space grows much larger than the one
typically used in the Bayesian preference elicitation literature, and
as such represents a good testbed for comparing the scalability of the
various methods. The feasible configuration space was encoded in {\sc
  SetMargin} through appropriate MILP constraints, while the other
methods require all datasets to be explicitly grounded.  Users were
simulated by drawing $20$ random utility vectors from each of four
different distributions. The first two mimic those used
in~\cite{guo2010real}\footnote{We used the parameters implemented in their code, which are slightly different from those reported in the paper.}%
%\footnote{The normal distribution reported
%  in~\cite{guo2010real} differs slightly from the one implemented in
%  the code of their experimental setting.  In our experiments we use
%  the latter.}
: (1) a uniform distribution over $[1, 100]$ for each
individual weight, and (2) a normal distribution with mean $25$
and %covariance
standard deviation $\frac{25}{3}$ (each attribute is sample i.i.d).
We further produced two novel {\em sparse} versions of the uniform and
normal distributions setting to zero $80\%$ of the entries (sampled
uniformly at random). We set a maximum budget of 100 iterations for all
methods for simplicity.

@@

%\paolo{any idea about possible future works?}
% FUTURE WORKS
